---
title: 'Reproducible Research: Peer Assessment 1'
author: "Pradeep Peddineni"
date: "December 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading Data
Our first job here is to load data

```{r}
setwd("C:/Users/ppeddi001c/Desktop/DataScience/R/activity")
activity<-read.csv('activity.csv')
head(activity)

```


## preprocessing the data
```{r}
library(dplyr)
totalSteps <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)

```

## What is mean total number of steps taken per day?


```{r}
 #caluclate mean and median
        mean_value = round(mean(totalSteps$steps,na.rm = TRUE), 1)
        median_value = round(median(totalSteps$steps,na.rm = TRUE), 1)

library(ggplot2)
a <- ggplot(totalSteps,show_guide=T,aes(x = steps))+ggtitle("Histogram of daily steps") +
            xlab("Steps (binwidth 2000)") +
            geom_histogram(binwidth = 2000)
b<-a+geom_vline(xintercept = median_value,lwd = 5, col = 'red')+geom_vline(xintercept = mean_value,lwd = 2, col = 'blue')

b

```

here i have tried to achieve having an legend with those mean and median variable values in ggplot but couldnt succeed however in a base hist plot i could. if u have any ideas please reply me below.

https://www.coursera.org/learn/reproducible-research/peer/gYyPt/course-project-1/discussions

```{r}

#So instead i will manually call mean and median values

mean(totalSteps$steps,na.rm = TRUE)


median(totalSteps$steps,na.rm = TRUE)

```

##What is the average daily activity pattern?



```{r}

#time series plot of the 5-minute interval 

stepsInterval <- aggregate(steps ~ interval, data = activity, FUN = mean, na.rm = TRUE)

names(stepsInterval) <- c("interval","steps")

stepline <- ggplot(stepsInterval,aes(interval,steps)) +
                 ggtitle("Time Series Plot of Average Steps by Interval") +
                 geom_line()
stepline  



```
```{r}
#The 5-min time interval contains the maximum number of steps?
stepsInterval[which.max(stepsInterval$steps),c("interval")]
```

##Imputing missing values

```{r}
#sub1 
#to Calculate total number of missing values in the dataset 

# total records
nrow(activity)

# missing records
sum(is.na(activity$steps))
```

```{r}
#sub2
#Replacing Missing Values 

library(plyr)
imputemean <- function(activity) replace(activity, is.na(activity), mean(activity, na.rm = TRUE))
```

```{r}
# Sub3
# Creating a New Dataset
activityimpute <- ddply(activity, ~interval, transform, steps = imputemean(steps))
```

```{r}
#sub4
#Histogram, Mean and Median

stepsdate <- aggregate(x = activityimpute$steps , by = list(activityimpute$date), FUN = sum ,na.rm=TRUE)

names(stepsdate) <- c("date","steps")

histplot <- ggplot(stepsdate,aes(x = steps)) +
            ggtitle("Histogram of daily steps after imputation") +
            xlab("Steps (binwidth 2000)") +
            geom_histogram(binwidth = 2000)
histplot 
```

```{r}

#New Mean and median for values with out nulls

mean(stepsdate$steps,na.rm = TRUE)


median(stepsdate$steps,na.rm = TRUE)
```

I do see the median has been increased little bit but mean remained the same.

##Are there differences in activity patterns between weekdays and weekends?

```{r}
#sub1
#New Factor Variable

activityimpute$dateP <- as.POSIXlt(activityimpute$date,format="%Y-%m-%d")
activityimpute$day <- "Weekday"
activityimpute$day [weekdays(activityimpute$dateP) %in% c("Saturday","Sunday")] <- "Weekend"


```

```{r}
#sub2
#Weekday/Weekend Comparison
finalactivityimpute <- aggregate(activityimpute[1],by=activityimpute[c(3,5)],FUN=mean,na.rm=TRUE)

library(ggplot2)
plot <- ggplot(data = finalactivityimpute,aes(x=interval,y=steps))
plot + geom_line() + facet_wrap(~day,nrow=2)


```


Here i did have some issue using "keep_md: true" statement
Error in yaml::yaml.load(enc2utf8(string), ...) : Scanner error: mapping values are not allowed in this context at line 4, column 23 Calls: <Anonymous> ... yaml_load_utf8 -> mark_utf8 -> <Anonymous> -> .Call Execution halted

So instead i have used the render func to get the .md file.

 library(rmarkdown)
 render("PA1_template.Rmd", md_document())

##EOD







